Bayesian exponential random graph modeling of KWL Burial network

### 2020-06-16 analysis/paper/paper.Rmd
some edits and feedback on methods section

## Bayesian modeling 
<!--Bayesian inference for Exponential Random Graph Models--> 
Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019]. This provides better estimations for complex social network models with heterogeneous data [@Caimo2017]. For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced by using Bayesian modelling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
- weighted data

@@

We analyze archaeological burial networks from a Iron Age site, Kiwulan, in northeastern Taiwan. We use burial data collected from the published excavation report, and original fieldwork note with detailed descriptions about context and stratigraphy. Given that burials in Chinese phase are less than previous two phases, we only compare the quantity and types of burial goods in burials in Chinese phase rather than network analysis. For the quantity of burial goods, we assigned the burials into three classes, high, medium, and low as an index of wealth as a variable to indicate the extent of the accumulation of wealth. In addition to wealth, we also consider another two social variables, age and gender, serving as indicators for distinguishing ascribed or achieved status in a society [citation].

The purpose of our analysis is to understand the underlying mechanisms of the network structures and how their relationships have an influence on the accumulation of wealth between different foreign contact periods. We hypothesize that after the arrival of the European presence in the 17th century, a group of ambitious individuals competed with each other for trading with the Europeans due to the high value of trade goods in their society, which led to the accumulation of high quality trade goods. Social inequality gradually increased among Kiwulan society, which can be observed from the burials from a more equal distribution of grave goods to a more differentiation in goods accumulation. Thus, we predict that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects in burial network after the European presence compared to the network before the European presence.

@@

We use Bayesian reference on Exponential Random Graph Models (ERGMs) to conduct statistical network analysis to test what variables may have contributed to the economic inequality observed between burials over time under foreign contact conditions. ERGMs assume that each network tie is a random variable which is dependent to each other somehow through the shared attributes or a reciprocal relationship [@Robins2007]. The dependence of ties can be hypothesized by parameters correspond to configurations in the network, such as a single tie, or a transitive triad. A model is built up by a distribution of random graphs according to these configurations that are also called network statistics. According to the formation of ties, network statistics can be divided into two ways, endogenous statistics and exogenous statistics [@Amati2019]. In endogenous model, the dependence of ties is decided by the presence of other ties, while in exogenous model, the dependence of ties is determined by node attributes. We use both exogenous and endogenous statistics to understand the characteristics of burial network and tendency of burials correlate with each other. To describe the relationship linked by shared burial goods better, the network is weighted to capture the multiple effects [@Snijders2006]. The network variables we examine include connectivity, clusters, centralization, and homophily in a undirected relations to reflect the social relations between individuals guided by social structures. Table \@ref(tab:ergm-terms-table) lists the network statistics we used in the model construction for burials with archaeological interpretations. Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].

@@

We conduct network analysis using R programming with the ergm package for simulation and model specification [@Hunter2008; @Morris2008], and then bergm package for Bayesian modeling [@Caimo2014].
We set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10000. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with a set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.

@@

>>>>

## Bayesian modeling

<!--Bayesian inference for Exponential Random Graph Models--> 
Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019]. This provides better estimations for complex social network models with heterogeneous data [@Caimo2017]. For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced by using Bayesian modelling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
- weighted data

@@

We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan to generate networks. We used burial data collected from the published excavation report, and original fieldwork notes with detailed descriptions about context and stratigraphy. We do not include the burials in Chinese phase in our network analysis because of the small sample size. <!-- but we will briefly discuss these in the discussion section --> We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods. In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].

The purpose of our analysis is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships. We hypothesize that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society. This led to the accumulation of high quality trade goods among a small numbers of individuals. Social inequality gradually increased in Kiwulan, and we predict this can be observed from the burials. We hypothesize a change from a more equal distribution of grave goods to a more differentiated distribution of goods. Thus, we predict that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects <!-- the reader needs to be introduced to these terms before you can use them like this... instead, need to foreground or foreshadow these concepts... "We used network analysis statistics to quantify the relationships and formally test our hypothesis. "--> in burial network after the European presence, compared to the network before European presence.

@@

We use Bayesian reference on Exponential Random Graph Models (ERGMs) to conduct statistical network analysis to test what variables may have contributed to the economic inequality observed between burials over time under foreign contact conditions.<!-- seems like a long sentence --> ERGMs assume that each network tie <!-- does the reader know what this is?--> is a random variable which is dependent on each other somehow through the shared attributes or a reciprocal relationship [@Robins2007]. The dependence of ties can be hypothesized by parameters corresponding to configurations in the network, such as a single tie, or a transitive triad <!-- does the reader know what is 'transitive? --> . A model is built up by a distribution of random graphs according to these configurations that are also called network statistics. According to the formation of ties, network statistics can be divided into two types, endogenous statistics and exogenous statistics [@Amati2019]. In endogenous models <!--statistic or model? what exactly? these seem like two different things that are getting mixed up. Choose one way to describe and stick to it to avoid confusing the reader -->, the dependence of ties is decided by the presence of other ties, while in an exogenous model, the dependence of ties is determined by node attributes <!-- node? -->. We use both exogenous and endogenous statistics to understand the characteristics of the burial network at Kiwulan and the correlate of burials. To describe the relationship linked by shared burial goods, the network is weighted to capture the multiple effects <!-- of what? the reader doesn't know about any effects yet --> [@Snijders2006]. The network variables we examine include connectivity, clusters, centralization, and homophily in undirected relations <!-- something wrong here, 'relations' seems wrong here --> to reflect the social relations <!-- avoid repeating the same work in a sentence --> between individuals guided by social structures. Table \@ref(tab:ergm-terms-table) lists the network statistics we used in the model construction for burials with archaeological interpretations <!-- put this earlier in the paragraph --> . Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].

@@

We conducted a network analysis using the R programming language with the ergm package for simulation and model specification [@Hunter2008; @Morris2008]. We used the bergm package for Bayesian modeling [@Caimo2014]. <!-- why did we use two packages? why not just one? this needs to be expanded a bit to make it clear to the reader --> We set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.

@@

<!-- "We test our hypothesis by measuring reciprocality,  transitivity, and  popularity structure effects and expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects" This is where you can connect the anthropological question, the archaeological hypothesis and the statistical method for testing the hypothesis  -->
